---
title: "Reproducible Workflow"
author: "by Lorraine Gaudio"
date:   "`r paste('Activity generated on', format(Sys.Date(), '%B %d, %Y'))`"
team: "Fall 2026"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Class Activity (Part One) 

A reproducible workflow enables someone else to run your analysis on a different computer and obtain the same results. Today, you will package your data, containing minimal transforms, into a small, portable project using the appropriate file formats for the target audiences (R users vs. non-R users). This activity is offered as an optional opportunity to earn technical skill points.

**Lesson Overview**

By the end of this activity you will be able to:

1. ⚡ Remember – common save formats and their trade-offs.

2. 🔍 Understand – why Projects + relative paths support portability.

3. 📍 Apply – a minimal Lesson-6 transform and save results to data_clean/.

4. 📊 Analyze – object structure after saving and re-loading.

5. 🛠️ Evaluate – which format to share with R vs. non-R users.

6. 🧰 Create – a runnable, documented mini-pipeline that another student can execute in Part 2.

Keep these goals in mind as you move through each section.

## What you'll do

**Part one** (40 min): 

1. Create a new RStudio Project and folders (optional)

2. Load a raw dataset (built-in or your own CSV)

3. Apply a minimal transform (e.g., from Lesson 6)

4. Save the cleaned dataset as .rds (for R users) and .csv (for non-R users)

**Part two** (preview): 

We'll swap your .rds (and .csv) + 01_save_data.Rmd with a peer. Your peer will clone your structure, re-run your script, load your .rds, and verify that object names, dimensions. They’ll document any issues and fixes and offer suggestion on other minimal transformations. (No action needed now.)

1. Load your peer's .rds to verify it works

2. Explore the object structure

3. Apply a minimal transform (e.g., from Lesson 6-7)

4. Save the cleaned dataset as .rds (for R users) and .csv (for non-R users)

## Points Rubric

This optional activity could earn you up to 2 technical skill points!

| Task | Points |
|-------|--------|
| Create a new RStudio Project and folders | 0.0 |
| Load a raw dataset (built-in or your own CSV) | 0.25 |
| Apply a minimal transform (e.g., from Lesson 6) | 1.50 |
| Save the cleaned dataset as .rds (for R users) and .csv (for
 non-R users) | 0.5 |
| Total | 2 |

## ❓ What is an .rds?

It’s a single, compressed R object saved to disk (serialization). It preserves the object's classes, factors, dates, list-columns, attributes. This allows you to share data and workflows with collaborators. The purpose of this activity is to practice saving a data frame from your R session as different kinds of shareable files. This is an essential step in a reproducible R workflow.

See this quick format guide for when to use .rds, .RData, .csv, and .xlsx.

1) .rds

- Use when: you want a single object saved exactly as-is, for fast, reliable re-use inside R projects.

- Pros: preserves types/attributes; fast; compact; explicit (you assign the name on load).

- Cons: R-specific; not human-readable; one object per file.

2) .RData (created with `save()`/`load()`)

- Use when: you want to snapshot many objects together.

- Pros: multiple objects; same fidelity as `.rds`.

- Cons: `load()` dumps objects into your workspace by their old names (less explicit/reproducible); R-only.

3) .csv

- Use when: you need a simple, universal text file to share with anyone (Python, Excel, Stata, etc.).

- Pros: human-readable; portable; tiny dependency surface.

- Cons: loses metadata (factors, classes); can mangle types (e.g., character vs numeric); can’t store list-columns; larger than .rds for complex data.

4) .xlsx (Excel workbook)

- Use when: a stakeholder requires Excel (multi-sheet handoff, lightweight manual review).

- Pros: multi-sheet; familiar to non-coders.

- Cons: manual edits reduce reproducibility; type guessing can break things (dates, leading zeros); heavier dependencies; not ideal for scripted pipelines.

**Rule of thumb**: use .rds for your internal, scripted workflow; export .csv (or .xlsx only if asked) for sharing with *non-R users*; use .RData sparingly for multi-object checkpoints.

# Mini reproducible workflow

Below is a starter pattern you can run on any machine with R + RStudio.

## Setup (5 min)

Create a new RStudio Project and folders (optional), move your raw data to the correct folder, and create a new R Markdown file.

### Make a Project & folders

Step one: Create a new RStudio Project (File > New Project > New Directory > New Project).

Step two: Name it something relevant (e.g., "`save_rds_activity`") within the folder of this course. This will create a new folder nested inside your course folder and a new .Rproj file.

Step three: 🎯 Create folders: data, scripts, output.

```{r}
# ⚡ Create folders
dirs <- c("data_raw", "data_clean")
invisible(lapply(dirs, dir.create, showWarnings = FALSE))
```

✅ This creates folders, (`data_raw` and `data_clean`) if they don't already exist, and suppresses warnings if they do. 👀 Check your file explorer or finder to verify the folders are there. 

### R Markdown

Step one: Create a new R Markdown file (File > New File > R Markdown).

Step two: Name it and save it as `01_save_data.Rmd`. Save it in the Project root (same level as data_clean/).

## Load Raw data (5 min)

Choose ONE path

You need to choose a "raw" dataset for this activity. You can use (A) a built-in/package dataset or (B) a CSV on your computer (place it in data_raw/). 

Ideally, you will use the dataset you will be using for your signature assignment. If you do not have a dataset ready for your signature assignment, consider using a dataset from dslabs for this activity.  

```{r}
# Option A — use a package dataset
# dslabs package
library(dslabs)  # For more datasets
utils::data(package = "dslabs")$results[, "Item"]
```

🎯 If you choose a dataset from dslabs, you can load it with data("dataset_name").

```{r}
# ⚡ load dataset from dslabs
data("movielens")
# ✅  Verify
head(movielens)
```

🎯 I use `mtcars` for an example in this activity of how to use a built-in dataset. 

```{r}
# ⚡ Example "import" step (pretend mtcars is your raw data)
mtcars_df <- mtcars
# ✅  Verify
mtcars_df[1:6 , c(1, 4)]   
```

If you choose a CSV dataset from your own computer, read it in with `read.csv()` or `readr::read_csv()`. Place the raw data file in the data_raw folder you created above. As an example of loading and external dataset, 🎯 I'll load `signature_assignment_example.csv` from the `data_raw` folder.

```{r}
# Option B — use your own CSV (put it in data_raw/)
# ⚡ Example load dataset from your computer
example_df <- read.csv(file.path("data_raw", "signature_assignment_example.csv"))
# ✅
example_df[1:6 , c(1:2,19)]  # show first 6 rows and a few columns
```

🗣 The dataset loaded successfully.

## Clean/transform in scripts (15 min)

Make a transformation or two to the dataset. Your example can do minimal transformations. Apply any steps you would like to practice from Lesson 6.

### mtcars example transformations

```{r}
# Step one: explore the data
mtcars_df[1:6 , c("mpg", "cyl")]

```


🎯 I add a new column called "model" that contains the row names. 

```{r}
# ⚡ keep model names as a column
mtcars_df[ , "model"] <- rownames(mtcars_df)
```

```{r}
# ✅ Verify:
mtcars_df[1:6 , c("model", "mpg", "cyl")]
```


🗣 The change was successful. rownames(mtcars_df) collects the rownames as a vertex then mtcars_df[ , "model"] assigns the vertex to a new column called "model".

🎯 I remove the row names.

```{r}
# ⚡ remove row names
rownames(mtcars_df) <- NULL

```

```{r}
# ✅ Verify: 
mtcars_df[1:6 , c("model", "mpg", "cyl")]
```

🗣 The change was successful. The rownames(mtcars_df) <- NULL command successfully removed the row names.

### Signature assignment example transformations

```{r}
# Step one: explore the data
example_df[1:6, c("id", "humans_versus_AI_control")]
```

🎯 I want to make two different types of columns for the AI Competency results section. I would like to have the numeric values in one column and the words in another column. This will make it easier to analyze.

```{r}
# What are my options
unique(example_df$humans_versus_AI_control) # An example column
```


🎯 Using ifelse() and overwriting values from lesson 6, I will create a new column called `humans_versus_AI_control_num` that contains numeric values for the `humans_versus_AI_control` column. I will then overwrite the values in the original column so that it only contains the words. 

1) 🎯 Create a numeric column with ifelse()

```{r}
# ⚡ Create numeric column for humans_versus_AI_control
example_df[ , "humans_versus_AI_control_num"] <-
  ifelse(example_df[ , "humans_versus_AI_control"] == 
           "1. No Awareness", 1L,
  ifelse(example_df[ , "humans_versus_AI_control"] == 
           "2. Awareness",    2L,
  ifelse(example_df[ , "humans_versus_AI_control"] == 
           "3. Aquire",       3L,
  ifelse(example_df[ , "humans_versus_AI_control"] == 
           "4. Deepen",       4L,
  ifelse(example_df[ , "humans_versus_AI_control"] == 
           "5. Create",       
         5L, NA_integer_)))))
```
```{r}
# ✅ Verify: 
example_df[1:6 , c("id", "humans_versus_AI_control", "humans_versus_AI_control_num")]
```

🗣 A new column called `humans_versus_AI_control_num` was successfully created.

2) 🎯 Overwrite the original column to keep only the words.

```{r}
# ⚡ Overwrite original column to keep only the words
example_df[ example_df[ , "humans_versus_AI_control"] == 
              "1. No Awareness" , "humans_versus_AI_control"] <- "No Awareness"
example_df[ example_df[ , "humans_versus_AI_control"] == 
              "2. Awareness"    , "humans_versus_AI_control"] <- "Awareness"
example_df[ example_df[ , "humans_versus_AI_control"] == 
              "3. Aquire"       , "humans_versus_AI_control"] <- "Aquire"
example_df[ example_df[ , "humans_versus_AI_control"] == 
              "4. Deepen"       , "humans_versus_AI_control"] <- "Deepen"
example_df[ example_df[ , "humans_versus_AI_control"] == 
              "5. Create"       , "humans_versus_AI_control"] <- "Create"
```

```{r}
#✅ Verify: 
example_df[1:6 , c("id", "humans_versus_AI_control", "humans_versus_AI_control_num")]
```

3) 🎯 I select only the columns I want to keep for my final dataset to share.

```{r}
# ⚡ Select only the columns I want to keep
example_df_clean <- example_df[ , c("id", "humans_versus_AI_control", "humans_versus_AI_control_num")]
```

```{r}
# ✅ Verify: 
example_df_clean[1:6, ]
```

# Save cleaned data (5 min)

We’ll save the cleaned object twice: once as .rds (for R pipelines) and once as .csv (for sharing).

## RDS

🎯 `file.path()` builds correct paths on Windows/Mac/Linux and keeps your code portable inside an RStudio Project.

```{r}
# ⚡ Create a file save path
rds_path <- file.path("data_clean", "dataset_clean.rds")
```

This creates a path to save the cleaned dataset as `dataset_clean.rds` in the `data_clean` folder. You can change `dataset` to something relevant to your dataset.

```{r}
# ⚡/ ✅   
saveRDS(mtcars_df, rds_path) # change mtcars_df to the name of your cleaned dataset
```
👀 Check your Files pane (RStudio) to verify the folders are there.


## CSV

```{r}
# ⚡ Create a file save path
csv_path <- file.path("data_clean", "dataset_clean.csv")
```

```{r}
# ✅ CSV for non-R users
write.csv(mtcars_df, csv_path)

```

## Submission (2 min)

Upload the folowing files to Module 6 → Class Activity 6.

- 01_save_data.Rmd (the R Markdown you used for this activity)

- data_clean/dataset_clean.rds

- data_clean/dataset_clean.csv

## Today you practiced: 

- worked inside an RStudio Project with relative paths,

- loaded a raw dataset (package or CSV),

- performed a minimal, documented transform (Lesson 6 skill),

- saved a clean dataset for two audiences: .rds for R users and .csv for non-R users,

- verified your results by reloading/checking structure.

This gives you a portable, re-runnable mini-pipeline that your collaborator can execute in Part Two.


### P.S. How to load rds


```{r, eval=FALSE}
# ✅ 
cars <- readRDS("mtcars_clean.rds")  # adjust file path as needed
```


