---
title: "🤪 Vector MadLib"
author: "by Lorraine Gaudio"
date:   "`r paste('Version', format(Sys.Date(), '%B %d, %Y'))`"
team: "Summer 2026"
output: 
  html_document: # To create an HTML document from R Markdown
    toc: false # Table of contents (TOC)
    toc_depth: 1 #(meaning that level 1, 2, and 3 headers will be included in the table of contents
    toc_float: # Float the table of contents to the left of the main document
      collapsed: false # Collapsed (defaults to TRUE) controls whether the TOC appears with only the top-level
      smooth_scroll: true # controls whether page scrolls are animated when TOC items are navigated to via mouse clicks.
    number_sections: true # Numbering starts with "#" (H1). Without H1 headers, the H2 headers ("##") will be numbered with 0.1, 0.2, and so on.
    css: ../assets/styles.css # This is the name of the CSS file to style the HTML document with Boise State Brand. The CSS file must be in the same directory as the R Markdown file.
    fig_caption: true #Whether figures are rendered with captions.
    df_print: paged # Printing data frames with interactive scrolling
    code_folding: show # Enables you to include R code but have it hidden by default. (Show hide button)
    includes:
      in_header: ../assets/header.html
      after_body: ../assets/footer.html
encoding: UTF-8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=FALSE, include=FALSE}
title: "Chapter 3 Activity"
author: "by Lorraine Gaudio"
date:   "`r paste('Lesson generated on', format(Sys.Date(), '%B %d, %Y'))`"
output: 
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
    citation_package: natbib
    fig_caption: true
    df_print: kable # Data frame printing
    includes:
      in_header: ../assets/header.tex
    latex_engine: xelatex  # Use xelatex to support fontspec
fontsize: 12pt
geometry: margin=1in
mainfont: "Garamond" # Sets the font of the entire document
sansfont: "Gotham-Book.otf" # Set sans-serif font to Gotham Book
monofont: "Courier New" # Set monospace font to Courier New
documentclass: scrreprt
linkcolor: boisestateblue # Customizes the color of hyperlinks
urlcolor: magenta # Customizes the color of URLs
citecolor: black # Customizes the color of citations
bibliography: references.bib # Bibliography file
biblio-style: apalike                 # ⟵ natbib needs a .bst style
natbiboptions: "round,authoryear"     # round brackets, Author (Year)
 
Or
title: "Chapter 3 Activity"
author: "by Lorraine Gaudio"
date:   "`r paste('Version', format(Sys.Date(), '%B %d, %Y'))`"
team: "SPring 2026"
output: 
  html_document: # To create an HTML document from R Markdown
    toc: true # Table of contents (TOC)
    toc_depth: 1 #(meaning that level 1, 2, and 3 headers will be included in the table of contents
    toc_float: # Float the table of contents to the left of the main document
      collapsed: false # Collapsed (defaults to TRUE) controls whether the TOC appears with only the top-level
      smooth_scroll: true # controls whether page scrolls are animated when TOC items are navigated to via mouse clicks.
    number_sections: true # Numbering starts with "#" (H1). Without H1 headers, the H2 headers ("##") will be numbered with 0.1, 0.2, and so on.
    css: ../assets/styles.css # This is the name of the CSS file to style the HTML document with Boise State Brand. The CSS file must be in the same directory as the R Markdown file.
    fig_caption: true #Whether figures are rendered with captions.
    df_print: paged # Printing data frames with interactive scrolling
    code_folding: show # Enables you to include R code but have it hidden by default. (Show hide button)
    includes:
      in_header: ../assets/header.html
      after_body: ../assets/footer.html
```


# Overview

This coding challenge is designed to deepen your understanding of vectors in R. You will build vectors of different types, convert one existing character vector into a factor, and generate two MadLib stories from your objects.


In this activity, you will practice how to:

1. 🔨 Build and inspect vectors: numeric, integer, logical, and character

2.  🔧 Use logical operators to create logical vectors

3. 🪄 Convert a character vector into factor class

4. 💪 Practice reproducible work by fixing errors and documenting what happened

## How to complete this activity

Open the provided R Notebook template for this activity. The template already contains the main sections, starter code chunks, and places for your comments/memos.

Your job is to work through the template from top to bottom:

1. Fill in the blanks in the code chunks.
2. Run each chunk after you complete it.
3. Check that the required objects appear in your Environment.
4. Write short comments or memos when prompted.
5. Run the story code after all required objects have been created.

💡 **Tip:** The story code depends on the vectors you create earlier in the notebook. If an object is missing, misspelled, or created with the wrong name, the story code will break.

# Vectors

**Rules**: Name your vectors *exactly* as listed. Follow the rules for each type. Element duplication is allowed. 

## Numeric vector

🎯  Create a vector called `numeric` using your pick of six **real numbers** (length = 6)

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
numeric <- c( __ , __ , __ , __ , __ , __) # length = 6
```

```{r eval=FALSE}
## ✅  print stored vector
numeric
```

🎯  Use `length()` to verify that `numeric` contains six values.

```{r eval=FALSE}
## ✅ Check the length of your vector
_____(numeric)
```

🗣 Comment: Describe what kind of values are stored in the **numeric vector**.

## Integer vector

🎯 Create a vector called `integer` using four **whole numbers**, your pick. (length = 4)

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
integer <- c(__L, __L, __L, __L) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
integer
```

🎯  Use `length()` to verify that `integer` contains exactly four values.

```{r eval=FALSE}
## ✅ Check the length of your vector
_____(integer)
```

🗣 Comment: Describe what makes this vector an **integer vector**.

## Logical vector

A logical vector can be created several ways. For this task, you will create the vector using **comparison operators**.

___

📖 **Review** The comparison operators are:

- `>` greater than

- `<` less than

- `==` equal to

- `>=` greater than or equal to

- `<=` less than or equal to

- `!=` not equal to


___

**EXAMPLE**

```{r, eval=FALSE}
logical_example <- c(1 > 2, (3 + 5) <= (10 - 2)) # length = 2
```


___

🎯  Create a vector called `logical` using only **logical operators** (length = 4).

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
logical <- c( __ , __ , __ , __ ) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
logical
```

🎯  Use `class()` to verify that the object `logical` has the class `logical`. Use `length()` to verify there are four elements in the `logical` object.

```{r eval=FALSE}
## ✅ Check the length and class of your vector
_____(logical)
_____(logical)
```

🗣 Comment: Explain one comparison you used and why it returns `TRUE` or `FALSE`.

## Character vectors

Make each character vector with words or phrases of your choice. 💡 **Tip:** character values need quotation marks.

### Name Vector

🎯  Create a vector called `name`. Use any people names, real or fictional (length = 6); Element duplication is recommended.

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
name <- c( __ , __ , __ , __ , __ , __ ) # length = 6
```

```{r eval=FALSE}
## ✅  print stored vector
name
```

### Team Name

🎯  Create a vector called `group_name` → imaginary team/group names (length = 4)

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
group_name <- c( __ , __ , __ , __ ) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
group_name
```

### Social Handle

🎯  Create a vector called `at_handle` → any made-up social media handles (length = 4)

_A social media handle is a unique identifier on a platform. Prefixed with “@”, it is used to find, tag, or mention someone, group, or organization online._

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
at_handle <- c( __ , __ , __ , __ ) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
at_handle
```

### Item

🎯  Create a vector called `item` → any kind of thing or object (length = 6). Element duplication is recommended.

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
item <- c( __ , __ , __ , __ , __ , __ ) # length = 6
```

```{r eval=FALSE}
## ✅  print stored vector
item
```

### Activity

🎯  Create a vector called `activity` → something you do (length = 4).

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
activity <- c( __ , __ , __ , __ ) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
activity
```

### Location

🎯  Create a vector called `location` → any place, real or imagined (length = 4)

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
location <- c( __ , __ , __ , __ ) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
location
```

### Rally Cry

🎯  Create a vector called `call_to_action` → a catch phrase or statement (length = 4)

_A rally cry is a phrase, or idea used to inspire, unite, or motivate a group toward a common goal or action._

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
call_to_action <- c(__, __, __, __) # length = 4
```

```{r eval=FALSE}
## ✅  print stored vector
call_to_action
```

### Emotion

🎯  Create a vector called `emotion` → a mood or feeling (length = 6). Element duplication is recommended.

```{r eval=FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
emotion <- c(__, __, __, __, __, __)
```

```{r eval=FALSE}
## ✅  print stored vector
emotion
```

### Retro Game

🎯  Create a vector called `retro_game` → any game titles (length = 4)

```{r eval = FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
retro_game <- c(__, __, __, __)
```

```{r eval=FALSE}
## ✅  print stored vector
retro_game
```

### Social Media

🎯  Create a vector called `social_media_platform` → any virtual community  (length = 3)

_A social media platform is a website or application that allows users to create and share content or participate in social networking._

```{r eval = FALSE}
# ⚡ Type this in your R Notebook, Fill in the blanks
social_media_platform <- c(__, __, __) # length 3
```

### Character Vector Check

```{r eval=FALSE}
# ✅ Check lengths
length(name)
length(group_name)
length(at_handle)
length(item)
length(activity)
length(location)
length(call_to_action)
length(emotion)
length(retro_game)
length(social_media_platform)
```

🗣 Comment: Describe one issue that can happen if character values are missing quotation marks.

## Factor class

**Create a factor for story energy**

A factor is R's way of storing categorical data. Factors often print like words, but R stores them internally using category labels called **levels**.

Factors matter when categories have a meaningful order.

For example, the story energy categories below work like a simple rating scale:

`low` → `medium` → `high`

That order makes sense to a human. But if R stores the values as plain character strings, R may display them alphabetically:

`high` → `low` → `medium`

That is not the order we want.

Later in data analysis, factors help R display tables, summaries, and plots in the order that makes sense for the data.

### Step 1

**Create a character vector**

🎯 Create a character vector called `story_energy_raw` with six values. Use only `"low"`, `"medium"`, and `"high"`. 💡 **Tip:** You will need to repeat values.

```{r, eval=FALSE}
# ⚡ Create a character vector of story energy categories
story_energy_raw <- c(__, __, __, __, __, __)
```

```{r, eval=FALSE}
# ✅ Check the original class
class(story_energy_raw)
```

🗣 Comment: What class is `story_energy_raw` before conversion?

```{r, eval=FALSE}
# ✅ Check the original class
class(story_energy_raw)
```

🗣 Comment: What class is `story_energy_raw` before conversion?

```{r, eval=FALSE}
# ✅ See how R displays the categories before conversion
table(story_energy_raw)
```

🗣 Comment: What order does R use when it displays the character values in the table?

### Step 2

**Convert to factor class**

🎯 Convert `story_energy_raw` into a factor called `story_energy`. Levels are set in a meaningful order.

```{r eval=FALSE}
# ⚡ Convert to factor and control the level order
story_energy <- factor(
  story_energy_raw,
  levels = c("low", "medium", "high")
)
```

### Step 3

**Check the factor**

🎯 Verify that `story_energy` has factor class and inspect its levels.

```{r eval=FALSE}
# ✅ Check the class and levels
class(story_energy)
levels(story_energy)
```
```{r eval=FALSE}
# ✅ See how R displays the categories after conversion
table(story_energy)
```

🗣 Comment: What changed after converting `story_energy_raw` into a factor?

🗣 Comment: What does it mean that the levels are in a "meaningful order" for this factor? 

🗣 Comment: Why might this matter later when creating a table, summary, or plot with this factor?

# Story

➤  Copy and paste the following code and read your Mad Lib Story!

➤ Before running the story code, confirm these objects exist in your Environment:

`numeric`, `integer`, `logical`, `name`, `group_name`, `at_handle`, `item`, `activity`, `location`, `call_to_action`, `emotion`, `retro_game`, `social_media_platform`, and `story_factor`

```{r , eval=FALSE}
# ⚡  Copy and paste this code into your R Notebook
# Function to pick a random element from a vector
pick <- function(x) sample(x, 1)

# ---------- Story 1 ----------
story1 <- paste(
  "After a long week,", pick(name), "unplugged by trying", pick(activity),
  "with", pick(group_name), "at their favorite spot,", paste0(pick(location), "."),
  "Vibe rating:", paste0(pick(numeric), "."),
  "The ", pick(activity), "was going so well, but then a crucial missing", paste0(pick(item), " was discovered."),
  "Going with the flow was", ifelse(pick(logical), "reasonable.", "unreasonable."),
  "Reminding each other", paste0(pick(call_to_action), "."),
  "They made a plan for Episode", paste0(pick(integer), ","),
  "closing with a", paste0(pick(emotion), " mood.")
)

print(story1)
# ---------- Story 2 ----------
# Using the factor object to pick one category
selected_energy <- pick(story_energy)

story2 <- paste(
  "On a mission for a fashion comeback,", pick(name), "found", pick(item),
  "at", pick(location), "that screamed Y2K.",
  "They spent a total of", paste0(pick(numeric), " dollars"),
  "but they", ifelse(pick(logical), "remembered", "forgot"), "to get a receipt!",
  "After posting a 'before' shot on", paste0(pick(social_media_platform), ","),
  "the post went viral", ifelse(pick(logical), "by midnight.", "by the next day."),
  "The official story energy level was", paste0(selected_energy, "."),
  "Their retro level was officially", paste0(pick(retro_game), " status,"),
  "and they immediately upcycled the item online for", paste0(pick(numeric), " dollars"),  
  "and sold in", pick(integer), "minutes.", 
  "The community response was so strong, they had this final message for their followers:", 
  paste0(pick(call_to_action), ".")
)

print(story2)
```


# Glossary

**Arguments**: Inputs that you provide to functions to customize their behavior. Arguments are specified within the parentheses of a function call and can modify how the function operates or what data it processes.

**Character Vector**: A vector that stores text strings, written in quotes, such as `"apple", "banana", "grapes"` or `"ID_001", "ID_002", "ID_003"`.

**Class** `class()`: A function that reports the object’s class, which tells R how to interpret the object and which methods to use when functions are applied to it. For example, `class(3.14)` returns `"numeric"`, and `class(mtcars)` returns `"data.frame"`.

**Code Chunks**: Fenced blocks in an R Markdown/Notebook document where R code is written and executed, usually starting with <code>&#96;&#96;&#96;{r}</code>. Only code inside a chunk runs, and its output (including plots and tables) appears directly below the chunk when executed.

**Comparison Operators**: Operators that create or combine logical elements. The most common are comparison operators—`==`, `!=`, `<`, `>`, `<=`, `>=`—which return `TRUE`/`FALSE`.

**Concatenate** `c()`: A function that combines (concatenates) values into a single vector. You use `c()` to create a vector, append items to an existing vector, or join multiple vectors together (for example, `c(1, 2, 3)` or `c(x, y)`.

**Elements**: The individual data values stored inside a vector object. Each element has a position (its index) and can be accessed by indexing (for example, `x[1]` for the first element).

**Environment Pane**: The pane that lists the objects currently stored in memory for your session (for example, data frames, vectors, and functions). It helps you check what exists, inspect object names/types, and remove or manage objects.

**Factor Class**: An R object class for categorical data, where each value is stored as an integer code that points to a category label. Factors help R handle categories correctly in tables, plots, and statistical models, even though they may print like text.

**Functions**: Reusable blocks of code that perform specific tasks. Functions can take inputs (arguments) and return outputs. R has many built-in functions, and you can also create your own.

**Function Call**: The act of running a function by writing its name followed by parentheses, optionally supplying arguments inside (e.g., `mean(x)`). 

**Integer Vector**: A vector that stores whole numbers. In R, integers are typically written with an `L` suffix, such as `1L` or `42L`.

**Logical Vector**: A vector that stores Boolean elements: `TRUE` or `FALSE` (and sometimes `NA` for missing). Logical vectors are commonly produced by comparisons (for example, `x > 0`) and are heavily used for filtering and conditional code.

**Objects**: The variables, data frames, functions, and other items that you create and store in your R environment during a session. These objects can be viewed and managed in the Environment pane.

**Object Name**: The variable name you assign with `<-` that refers to the entire object in your environment. You use the object name to print, modify, or pass the whole object to functions.

**Numeric Vector**: A vector that stores real (decimal) numbers, such as `3.14`, `-2.5`, or `0`.

**Real Numbers**: Numbers that can be located on the number line, including whole numbers, integers, fractions, decimals, and irrational numbers such as `pi` and `sqrt(2)`. Real numbers do not include imaginary numbers, and in statistics they are commonly used to represent measured quantitative values such as height, weight, or time.

**Vector**: A one-dimensional object that stores an ordered sequence of elements of the same basic type (for example, all numeric, all character, or all logical). Vectors are the core building block of many R objects, and you create them with functions like `c()`; you can subset them by position or by logical conditions.

**Whole Numbers**: The counting numbers together with zero: `0, 1, 2, 3, ....` Whole numbers do not include negative numbers, fractions, or decimals.

 
