---
title: "🤪 Vector MadLib"
author: "YOUR NAME"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document: # To create an HTML document from R Notebook

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How to Use This Notebook

Work through this notebook from top to bottom.

1. Replace every blank `__` with your own values or code.
2. Run each code chunk after you complete it.
3. Do not rename the required objects.
4. Write short comments where prompted.
5. After all vectors are created, run the Story section.
6. If something errors, fix it and briefly document what happened.

# Vectors

A **Vector** is a one-dimensional object that stores an ordered sequence of elements of the same basic type (for example, all numeric, all character, or all logical). Vectors are the core building block of many R objects, and I create them with functions like `c()`; you can subset them by position or by logical conditions.


## Numeric vector

🎯  Create a vector called `numeric` using your pick of six **real numbers** (length = 6)

```{r}
# ⚡ 
numeric <- c( __ , __ , __ , __ , __ , __) # length = 6
```

```{r}
## ✅  print stored vector
numeric
```

🎯  Use `length()` to verify that `numeric` contains six values.

```{r}
## ✅ Check the length of your vector
_____(numeric)
```

🗣 Comment: Describe what kind of values are stored in the **numeric vector**.

> [Replace this text with your comment]


## Integer vector

🎯 Create a vector called `integer` using four **whole numbers**, your pick. (length = 4)

```{r}
# ⚡ 
integer <- c(__L, __L, __L, __L) # length = 4
```

```{r}
## ✅  print stored vector
integer
```

🎯  Use `length()` to verify that `integer` contains exactly four values.

```{r}
## ✅ Check the length of your vector
_____(integer)
```

🗣 Comment: Describe what makes this vector an **integer vector**.

> [Replace this text with your comment]

## Logical vector

A logical vector can be created several ways. For this task, you will create the vector using **comparison operators**.


🎯  Create a vector called `logical` using only **comparison operators** (length = 4).

```{r}
# ⚡
logical <- c( __ , __ , __ , __ ) # length = 4
```

```{r}
## ✅  print stored vector
logical
```

🎯  Use `class()` to verify that the object `logical` has the class `logical`. Use `length()` to verify there are four elements in the `logical` object.

```{r}
## ✅ Check the length and class of your vector
_____(logical)
_____(logical)
```

🗣 Comment: Explain one comparison you used and why it returns `TRUE` or `FALSE`.

> [Replace this text with your comment]

## Character vectors

Make each character vector with words or phrases of your choice. 💡 **Tip:** character values need quotation marks.

### Name Vector

🎯  Create a vector called `name`. Use any people names, real or fictional (length = 6); Element duplication is recommended.

```{r}
# ⚡ 
name <- c( __ , __ , __ , __ , __ , __ ) # length = 6
```

```{r}
## ✅  print stored vector
name
```

### Team Name

🎯  Create a vector called `group_name` → imaginary team/group names (length = 4)

```{r}
# ⚡ 
group_name <- c( __ , __ , __ , __ ) # length = 4
```

```{r}
## ✅  print stored vector
group_name
```

### Social Handle

🎯  Create a vector called `at_handle` → any made-up social media handles (length = 4)

```{r}
# ⚡ 
at_handle <- c( __ , __ , __ , __ ) # length = 4
```

```{r}
## ✅  print stored vector
at_handle
```

### Item

🎯  Create a vector called `item` → any kind of thing or object (length = 6). Element duplication is recommended.

```{r}
# ⚡ 
item <- c( __ , __ , __ , __ , __ , __ ) # length = 6
```

```{r}
## ✅  print stored vector
item
```

### Activity

🎯  Create a vector called `activity` → something you do (length = 4).

```{r}
# ⚡ 
activity <- c( __ , __ , __ , __ ) # length = 4
```

```{r}
## ✅  print stored vector
activity
```

### Location

🎯  Create a vector called `location` → any place, real or imagined (length = 4)

```{r}
# ⚡ 
location <- c( __ , __ , __ , __ ) # length = 4
```

```{r}
## ✅  print stored vector
location
```

### Rally Cry

🎯  Create a vector called `call_to_action` → a catch phrase or statement (length = 4)

```{r}
# ⚡
call_to_action <- c(__, __, __, __) # length = 4
```

```{r}
## ✅  print stored vector
call_to_action
```

### Emotion

🎯  Create a vector called `emotion` → a mood or feeling (length = 6). Element duplication is recommended.

```{r}
# ⚡ 
emotion <- c(__, __, __, __, __, __)
```

```{r}
## ✅  print stored vector
emotion
```

### Retro Game

🎯  Create a vector called `retro_game` → any game titles (length = 4)

```{r}
# ⚡
retro_game <- c(__, __, __, __)
```

```{r}
## ✅  print stored vector
retro_game
```

### Social Media

🎯  Create a vector called `social_media_platform` → any virtual community  (length = 3)

_A social media platform is a website or application that allows users to create and share content or participate in social networking._

```{r}
# ⚡
social_media_platform <- c(__, __, __) # length 3
```

### Character Vector Check

```{r}
# ✅ Check lengths
length(name)
length(group_name)
length(at_handle)
length(item)
length(activity)
length(location)
length(call_to_action)
length(emotion)
length(retro_game)
length(social_media_platform)
```

🗣 Comment: Describe one issue that can happen if character values are missing quotation marks.

> [Replace this text with your comment]

## Factor class

**Create a factor for story energy**

A factor is R's way of storing categorical data. Factors often print like words, but R stores them internally using category labels called **levels**.

Factors matter when categories have a meaningful order. For example, the story energy categories below work like a simple rating scale:

`low` → `medium` → `high`

That order makes sense to a human. But if R stores the values as plain character strings, R may display them alphabetically:

`high` → `low` → `medium`

That is not the order we want. Later in data analysis, factors help R display tables, summaries, and plots in the order that makes sense for the data.

### Step 1: Create a character vector

🎯 Create a character vector called `story_energy_raw` with six values. Use only `"low"`, `"medium"`, and `"high"`. 💡 **Tip:** You will need to repeat values.

```{r}
# ⚡ Create a character vector of story energy categories
story_energy_raw <- c(__, __, __, __, __, __)
```

```{r}
# ✅ Check the original class
class(story_energy_raw)
```

🗣 Comment: What class is `story_energy_raw` before conversion?

> [Replace this text with your comment]

```{r}
# ✅ See how R displays the categories before conversion
table(story_energy_raw)
```

🗣 Comment: What order does R use when it displays the character values in the table?

> [Replace this text with your comment]

### Step 2: Convert to factor class

🎯 Convert `story_energy_raw` into a factor called `story_energy`. Levels are set in a meaningful order.

```{r}
# ⚡ Convert to factor and control the level order
story_energy <- factor(
  story_energy_raw,
  levels = c("low", "medium", "high")
)
```

### Step 3: Check the factor

🎯 Verify that `story_energy` has factor class and inspect its levels.

```{r}
# ✅ Check the class and levels
class(story_energy)
levels(story_energy)
```

```{r}
# ✅ See how R displays the categories after conversion
table(story_energy)
```

🗣 Comment: What changed after converting `story_energy_raw` into a factor?

> [Replace this text with your comment]

🗣 Comment: What does it mean that the levels are in a "meaningful order" for this factor?

> [Replace this text with your comment]

🗣 Comment: Why might this matter later when creating a table, summary, or plot with this factor?

> [Replace this text with your comment]

# Story

Before running the story code, confirm these objects exist in your Environment:

`numeric`, `integer`, `logical`, `name`, `group_name`, `at_handle`, `item`, `activity`, `location`, `call_to_action`, `emotion`, `retro_game`, `social_media_platform`, `story_energy_raw`, and `story_energy`

```{r}
# Function to pick a random element from a vector
pick <- function(x) sample(x, 1)

# ---------- Story 1 ----------
story1 <- paste(
  "After a long week,", pick(name), "unplugged by trying", pick(activity),
  "with", pick(group_name), "at their favorite spot,", paste0(pick(location), "."),
  "Vibe rating:", paste0(pick(numeric), "."),
  "The ", pick(activity), "was going so well, but then a crucial missing", paste0(pick(item), " was discovered."),
  "Going with the flow was", ifelse(pick(logical), "reasonable.", "unreasonable."),
  "Reminding each other", paste0(pick(call_to_action), "."),
  "They made a plan for Episode", paste0(pick(integer), ","),
  "closing with a", paste0(pick(emotion), " mood.")
)

print(story1)

# ---------- Story 2 ----------
# Using the factor object to pick one story energy category
selected_energy <- pick(story_energy)

story2 <- paste(
  "On a mission for a fashion comeback,", pick(name), "found", pick(item),
  "at", pick(location), "that screamed Y2K.",
  "They spent a total of", paste0(pick(numeric), " dollars"),
  "but they", ifelse(pick(logical), "remembered", "forgot"), "to get a receipt!",
  "After posting a 'before' shot on", paste0(pick(social_media_platform), ","),
  "the post went viral", ifelse(pick(logical), "by midnight.", "by the next day."),
  "The official story energy level was", paste0(selected_energy, "."),
  "Their retro level was officially", paste0(pick(retro_game), " status,"),
  "and they immediately upcycled the item online for", paste0(pick(numeric), " dollars"),  
  "and sold in", pick(integer), "minutes.", 
  "The community response was so strong, they had this final message for their followers:", 
  paste0(pick(call_to_action), ".")
)

print(story2)
```

# Transparency Reflection

In 3–5 sentences, explain one thing you checked, fixed, or learned while completing this activity. If you used an outside resource, including AI, name the resource and explain what it helped you understand.

> Replace this text with your reflection.